Assessing the impact of urban heating on tree growth in Berlin with open inventory and environmental data

### 2022-05-14 analysis/paper/paper.Rmd
add AIC to results

The null models containing only age to explain diameter captured a large portion of deviance.
Including additional covariates thus resulted only in comparatively small increases of predictive skill.
The inclusion of temperature and temperature-age interactions improved the predictive skill, with models including the LandSat-derived UHI measures performing best on average.
Note that non-temperature models containing planting bed area and $LCZ6_{\frac{}{150~m}}$ alone outperformed several of those including other temperature measures and covariate combinations.
This indicated the importance of both covariates, with $LCZ6_{\frac{}{150~m}}$ also represented in the best-performing model, corroborating its impact on stem diameter development.
As $LCZ6_{\frac{}{300~m}}$ only marginally increased performance compared to the null model (age only), the reference scale for the impact of the urban fabric (i.e., LCZ6) likely has a limit at $<300~m$.
Nutrient availability, similar to $LCZ6_{\frac{}{300~m}}$, also only improved model performance marginally.
Building height, regardless of its reference scale (i.e., $150$ $vs$ $300~m$), as another proxy for urban fabric, showed minimal to no improvement compared to the age-only null model.

>>>>

The null models containing only age to explain diameter captured a large portion of deviance.
Including additional covariates thus resulted only in comparatively small increases of predictive skill.
The inclusion of temperature (additive structure) and temperature-age interactions improved the predictive skill, with interactions reaching higher explained deviance (i.e., skill).
Models including the LandSat-derived UHI measures performed best on average, where models containing $LCZ6_{\frac{}{150~m}}$ or all covariates (saturated model) did best (Figure$~$\@ref(fig:fig-model-deviance)), which also holds true for other temperature/UHI measures.
Note that non-temperature models containing planting bed area and $LCZ6_{\frac{}{150~m}}$ alone outperformed several of those including other temperature measures and covariate combinations.
This indicated the importance of both covariates, with $LCZ6_{\frac{}{150~m}}$ also represented in the best-performing model, corroborating its impact on stem diameter development.
As $LCZ6_{\frac{}{300~m}}$ only marginally increased performance compared to the null model (age only), the reference scale for the impact of the urban fabric (i.e., LCZ6) likely has a limit at $<300~m$.
Nutrient availability, similar to $LCZ6_{\frac{}{300~m}}$, also only improved model performance marginally.
Building height, regardless of its reference scale (i.e., $150$ $vs$ $300~m$), as another proxy for urban fabric, showed minimal to no improvement compared to the age-only null model.

```{r mod-aic, echo=FALSE, cache = TRUE}

mod_lcz6_tensor <- readRDS("./analysis/data/models/stat/filtered/mI_spatial_age_x_temp_by_species_lcz6_reBEZIRK_var-day_2007.Rds")
mod_lcz6_add <- readRDS("./analysis/data/models/stat/filtered/mI_spatial_age_ADD_temp_by_species_lcz6_reBEZIRK_var-day_2007.Rds")

mod_assessment <- rbind(broom::glance(mod_lcz6_add), broom::glance(mod_lcz6_tensor)) %>% 
  cbind(data.frame(mod = c("add", "tensor")))

aic_tens = mod_assessment$AIC[mod_assessment$mod == "tensor"]
aic_add = mod_assessment$AIC[mod_assessment$mod == "add"]

Differences in sample sizes between model structures due to covariate and/or temperature data coverage were considerable (e.g., planting bed area $vs.$ soil nutrient models for UrbClim UHI).
Thus, we considered the model with the overall greatest predictive skill and largest sample for further investigations, namely $LCZ6_{\frac{}{150~m}}$ for LandSat derived UHI.
For this variable combination, the model structure with age and temperature interaction also had a lower AIC ($AIC~=~8.95 \cdot 10^5$) compared to the additive one ($AIC~=~8.98 \cdot 10^5$).
<!-- For this variable combination, the model structure with age and temperature interaction also had a lower AIC ($AIC~=~`r format(aic_tens, digits = 3, scientific = TRUE)`$) compared to the additive one ($AIC~=~`r format(aic_add, digits = 3, scientific = TRUE)`$). -->
